### 2012-06-21 inst/examples/criticaltransition.Rmd
criticaltransition model

``` {r  echo=FALSE, cache=FALSE}
opts_knit$set(upload.fun = socialR::flickr.url)
orig = getOptions("device")
options(device = function(width = 5, height = 5) {
    pdf(NULL, width = width, height = height)
})
knitcitations::cleanbib()
````

# Reed Model

 * Author [Carl Boettiger](http://carlboettiger.info), <cboettig@gmail.com>
 * License: [CC0](http://creativecommons.org/publicdomain/zero/1.0/)
 * Description:  Implements a numerical version of the SDP described in `r knitcitations::citet("10.1016/0095-0696(79)90014-7")`.

@@

``` {r }
f <- function(x,h,p){
	S = x - h
	p[1] * S * (1 - S/p[2]) + S
}
````

With parameters `r` = `r r <- 1` and `K` = `r K <- 100`.

``` {r }
pars <- c(r, K)
````

@@

### Simulate 
Now we'll simulate 100 replicates of this stochastic process under the Reed optimal harvest policy determined above.

@@

``` {r cleanup}
options(device=orig)
````

>>>>

``` {r  echo=FALSE, cache=FALSE}
opts_knit$set(upload.fun = socialR::flickr.url)
orig = getOption("device")
options(device = function(width = 5, height = 5) {
    pdf(NULL, width = width, height = height)
})
knitcitations::cleanbib()
````

# Critical transition

 * Author [Carl Boettiger](http://carlboettiger.info), <cboettig@gmail.com>
 * License: [CC0](http://creativecommons.org/publicdomain/zero/1.0/)
 * Description:  Implements a numerical version of the SDP described in `r knitcitations::citet("10.1016/0095-0696(79)90014-7")`, in which the dynamics are slowly moving towards a tipping point.

@@

``` {r }
f <- May
````

With parameters

``` {r }
pars <- c(r = .75, k = 10, a=1, H=1, Q = 3)
K <- 8 # approx
````

@@

### Simulate

Now we'll simulate 100 replicates of this stochastic process, but with the system dynamics gradually changing over time.   
The bifurcation parameter will increase from `r pars[3]` to `r pars[3] + 2`.

``` {r simulationfunction}
ForwardSimulate <- 
  function(f, pars, x_grid, h_grid, x0, D, z_g,
         z_m=function(x) 1, z_i = function(x) 1, 
         profit=NULL){

  OptTime <- dim(D)[2]    # Stopping time
  x_h <- numeric(OptTime) # population dynamics with harvest
  h <- numeric(OptTime) # optimal havest level
  x_h[1] <- x0  # initial values
  s <- x_h # also track escapement
  x <- x_h # What would happen with no havest
  p <- numeric(OptTime)

  ## Simulate through time ##
  for(t in 1:(OptTime-1)){

    # Move towards collapse
    pars[3] <- pars[3] + 2/OptTime

    # Assess stock, with potential measurement error
    m_t <- x_h[t] * z_m()
    # Current state (is closest to which grid posititon) 
    St <- which.min(abs(x_grid - m_t)) 
    q_t <- h_grid[D[St, (t + 1) ]] 
    # Implement harvest/(effort) based on quota with noise 
    h[t] <- q_t * z_i()
    # Noise in growth 
    z <- z_g() 
    # population grows
    x_h[t+1] <- z * f(x_h[t], h[t], pars) # with havest
    s[t+1]   <- x_h[t] - q_t # anticipated escapement
    x[t+1]   <- z * f(x[t], 0, pars) # havest-free dynamics
    p[t] <- profit(x_h[t], h[t])
  }
  data.frame(time = 1:OptTime, fishstock = x_h, harvest = h,
             unharvested = x, escapement = s, profit = p) 
}

````

@@

``` {r cleanup, echo=FALSE, include=FALSE}
options(device='orig')
````
